---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cbiopourer

<!-- badges: start -->
<!-- badges: end -->

The goal of `cbiopourer `is to facilitate the creation of the files used for ingesting studies into cBioportal. It helps pour data into cBioportal.
A detailed explanation of the file formats can be found in https://docs.cbioportal.org/file-formats/#expression-data

## Installation

You can install the development version of cbiopourer from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("c1au6i0/cbiopourer")
```

## Use 

The meta_study, meta_clinical and the respective clinical data are required to load a study into the portal. The next sessions describe how to create those files using `cbiopourer`


### Meta Study

The meta cancer study (`meta_study.txt` contains metadata about the cancer study.

```{r}
library(cbiopourer)

# the Identifier for the study
cancer_study_identifier <- 
  
create_meta_study(
  folder_path = dir_out,
  type_of_cancer = "ptcls",
  cancer_study_identifier = cancer_study_identifier,
  name = "Peripheral T-cell Example",
  description = "Bulk RNA example"
)
```


