---
output: github_document
---

<!--README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cbiopourer

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/c1au6i0/cbiopourer/branch/main/graph/badge.svg)](https://app.codecov.io/gh/c1au6i0/cbiopourer?branch=main)
[![](https://img.shields.io/badge/devel%20version-0.0.0.99-blue.svg)](https://github.com/https://github.com/c1au6i0/cbiopourer)
<!-- badges: end -->

The purpose of cbiopourer is to simplify the process of generating files for importing studies into cBioportal. This package streamlines the transfer of data into cBioportal. For more information on the accepted file formats, please refer to https://docs.cbioportal.org/file-formats/#expression-data.

## Installation

You can install the development version of cbiopourer from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("c1au6i0/cbiopourer")
```

## Usage

In order to upload a study to the portal, the meta_study, meta_clinical, and corresponding clinical data files must be generated. The following sections explain how to use cbiopourer to create these files.

### Meta Study and Cancer Type

The meta_study.txt file contains metadata for the cancer study. The required arguments are illustrated in the example below. The function documentation details the other fields specified on the cBioportal website.

```{r}
library(cbiopourer)
library(dplyr)

# the Identifier for the study
cancer_study_identifier <- "example_study_2023"

# The directory where to create the file.
dir_out <- tempdir()
  
create_meta_study(
  folder_path = dir_out,
  type_of_cancer = "ptcls",
  cancer_study_identifier = cancer_study_identifier,
  add_global_case_list = "true",
  name = "Peripheral T-cell example",
  description = "Bulk RNA example",
  parent_type_of_cancer = "tissue"
)
```


The function generates a file in the format accepted by cBioportal.

```
type_of_cancer: ptcls
cancer_study_identifier: example_study_2023
name: Peripheral T-cell example
description: Bulk RNA example
add_global_case_list: true
```

If the type of cancer is not one of those already defined in cBioportal, a `meta-cancer_type.txt` file is required.

```{r}
create_cancer_type(
  folder_path = dir_out,
  type_of_cancer = "ptcls", # as listed in the meta-study.txt
  name = "Peripheral T-cell Lymphoma",
  dedicated_color = "HotPink",
  parent_type_of_cancer = "Lymphoma")
```

## Clinical files

According to cBioportal documentation, "clinical data is used to capture both clinical attributes and the mapping between patient and sample IDs". In most cases, you will need to provide information for both the patients and their corresponding samples, which requires the creation of metadata and data files (for a total of 4 files).

Let's suppose that our clinical information indicates the id of the patient and the relative cellline derived from the patient.

### Patients

```{r}
data(df_samples)
df_patients <- df_samples |> 
  dplyr::select(PATIENT_ID, CELLLINE) |> 
  dplyr::distinct(PATIENT_ID, .keep_all = TRUE)

glimpse(df_patients)
```

The following attributes regarding the 2 columns of `df_patients` are required to generate the appropriate files: 

- attribute Display Names.
- the attribute Descriptions.
- the attribute Datatype (STRING, NUMBER, BOOLEAN).
- the attribute Priority. 

We generate the info using.

```{r}
df_patients_datatype <- tribble(
  ~var_int, ~description, ~datatype, ~attr_priority, ~attr_name,
  "Patient ID", "patient id", "STRING", 1, "PATIENT_ID",
  "Cellline", "cellline id", "STRING", 1, "CELLLINE"
)
```

Now, we can supply those 2 datafames to the `create_clinical` function and create both the  metadata and data file. Note that since we are generating info regarding patients  the `datatype` argument of is set to "PATIENT_ATTRIBUTES".

**The function also checks that correct attributes are provided for each columns**.

```{r}
dir_out <- tempdir()
  create_clinical(
    folder_path = dir_out,
    cancer_study_identifier = cancer_study_identifier,
    datatype = "PATIENT_ATTRIBUTES", 
    clinical_meta = df_patients_datatype,
    clinical_dat = df_patients)
```

The clinical data will have a header formated as required by cBioportal.
```
#Patient<TAB>ID	Cellline<TAB>
#patient id<TAB>cellline id<TAB>
#STRING<TAB>STRING<TAB>
#1<TAB>1<TAB>
#PATIENT_ID<TAB>CELLLINE<TAB>
A<TAB>A<TAB>
B<TAB>B<TAB>
...
```

The metadata file generated is as follow:

```
cancer_study_identifier: example_study_2023
genetic_alteration_type: CLINICAL
datatype:  PATIENT_ATTRIBUTES
data_filename: data_clinical_patient.txt
```

### Samples

An example of sample info of a study is the following:

```{r}
data(df_samples)

glimpse(df_samples)
```
The steps to generate the files are the same described for the patient clinical info.

We start creating a file with required attributes of each of the columns of the `df_samples`.


```{r}
df_samples_datatype <- tribble(
  ~var_int, ~description, ~datatype, ~attr_priority, ~attr_name,
  "Sample ID", "sample id", "STRING", 1, "SAMPLE_ID",
  "Patient ID", "patient id", "STRING", 1, "PATIENT_ID",
  "Cancer Type", "cancer type", "STRING", 1, "CANCER_TYPE",
  "Cancer Type Detailed", "cancer type detailed", "STRING", 1, "CANCER_TYPE_DETAILED",
  "Cellline", "cellline id", "STRING", 1, "CELLLINE",
  "Time Treat", "time of administration of treatment", "STRING", 1, "TIME_TREAT",
  "Treatment", "drug admnistered", "STRING", 1, "TREAT"
)
```

We supply the 2 dataframes to the `create_clinical` function and we set the argument `datatype` to `SAMPLE_ATTRIBUTES`.

```{r}
create_clinical(
  folder_path = dir_out,
  cancer_study_identifier = cancer_study_identifier,
  datatype = "SAMPLE_ATTRIBUTES",
  clinical_meta = df_samples_datatype,
  clinical_dat = df_samples)
```

The clinical sample data file produced is as follow:

```
#Sample<TAB>ID<TAB>Patient<TAB>ID<TAB>Cancer<TAB>Type<TAB>CancerType Detailed<TAB>Cellline<TAB>Time Treat<TAB>Treatment
#sample id<TAB>patient id<TAB>cancer type<TAB>cancer type detailed<TAB>cellline id<TAB>time of administration of treatment<TAB>drug admnistered
#STRING<TAB>STRING<TAB>STRING<TAB>STRING<TAB>STRING<TAB>STRING<TAB>STRING
#1<TAB>1<TAB>1<TAB>1<TAB>1<TAB>1<TAB>1
#SAMPLE_ID<TAB>PATIENT_ID<TAB>CANCER_TYPE<TAB>CANCER_TYPE_DETAILED<TAB>CELLLINE<TAB>TIME_TREAT<TAB>TREAT
A_48h_VEH<TAB>A<TAB>Lymphoma<TAB>Peripheral T-cell Lymphoma<TAB>A<TAB>48h<TAB>VEH
A_48h_VEH<TAB>A<TAB>Lymphoma<TAB>Peripheral T-cell Lymphoma<TAB>A<TAB>48h<TAB>VEH
```

## Expression Counts
To create meta and data files for RNAsep expression we need a dataframe containing genes x sample counts and at least one coloumn named `Hugo_Symbol`, `Entrez_Gene_Id`. See details in https://docs.cbioportal.org/file-formats/#expression-data


```{r}
data("df_expr")
head(df_expr)
```

Few other info are required and are supplid to the `create_expression` function:

```{r}
create_expression(
  folder_path = dir_out,
  cancer_study_identifier = cancer_study_identifier,
  genetic_alteration_type = "MRNA_EXPRESSION",
  datatype = "CONTINUOUS",
  stable_id = "rna_seq_mrna",
  source_stable_id = NULL,
  show_profile_in_analysis_tab = FALSE,
  profile_name = "mRNA expression",
  profile_description = "Expression levels",
  df_expr = df_expr,
  gene_panel = NULL
)
```

